Gravitational Lens Recovery with \Glass: How to measure the mass profile and shape of a lens

### 2013-07-11 limits/ms.tex
editing intro

\cite{1981ApJ...244..736Y} were remarkably prescient about the
subsequent development of lens modelling.  First, the introduced the
technique of choosing a parametric form for the lensing mass and then
fitting for the parameters, which is still the most common strategy
\citep[see for example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.
Second, they pointed out the non-uniqueness of lens models, or lensing
degeneracies.  Third, they suggested combining lensing data with
stellar kinematics and X-rays, to reduce the effect of the
degeneracies.  Later work, as well as following up these suggestions,
has introduced some ideas.  Four of these are important in the present
work.
\begin{itemize}
\item In free-form models, there is no specified parametric form for
  the mass distribution.  There are still assumptions (or priors) on
  the mass distribution, such as smoothness or being centrally
  concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms.
  \cite{2006MNRAS.367.1209L} implemented a particular elegant prior,
  which is that the mass distribution must be non-negative and no
  extra images are produced.  Free-form are at present most common for
  cluster lenses\citep{2009ApJ...690..154S,2013arXiv1304.2393S}.
\item Model ensembles, exploring a diverse range of possible mass
  distributions that nonetheless all fit the data, are a way of
  cambating the non-uniqueness of models.  Model ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are
  more common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.
\item Stellar kinematics has played a role in many modelling
  techniques.  The underlying reason is the virial theorem, which
  relates the line of sight velocity dispersion to the Einstein radius
  as
\begin{equation}
  \frac{\langle v^2_{\rm los}\rangle}{c^2} \approx
  \frac{\theta_E}{4\pi} \frac{D_S}{D_{LS}}
\end{equation}
  with the relation becoming exact for isothermal lenses.  The use of
  two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is
  especially interesting.
\item The stellar mass in a lens can be inferred from photometry and
  compared with the total mass
  \citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.
  Since the inferred stellar mass does depend on the assumed IMF,
  lenses in which stellar mass dominates can be used to derive upper
  bounds on the stellar $M/L$ \cite{2010MNRAS.409L..30F}. Lower bounds
  on stellar $M/L$ have also been claimed \citep{2013MNRAS.428.3183D}
  but these require upper-bound assumptions about the dark-matter
  profile.
\item Testing modelling strategies to see how well they recover
  simulated lenses is increasingly recognized as essential.  Simple
  blind tests have appeared in earlier work \citep[for example, Figure
    2 in][]{2000AJ....119..439W}, but more recently, test against
  dynamically simulated galaxies are favoured
  \citep{2007ApJ...667..645R,2009MNRAS.393.1114B}.
\end{itemize}
There are also more modelling ideas in the literature.  One is to use
X-ray intensity and temperature profiles as a mass constraint
\citep[such as in][]{2013ApJ...765...25N}.  Another idea is to model
multiple lenses simultaneously, with one or more cosmological
parameters variable but shared between the lenses.  This strategy has
been used to constrain $H_0$ from time delays
\citep{2006ApJ...652L...5S,2008ApJ...679...17C,2010ApJ...712.1378P}.
In principle, cosmological parameters could be constrained even from a
single lens.  Some models have attempted this, starting from a prior
based on the values and uncertainties from CMB fluctuations
\citep{2010Sci...329..924J,2013arXiv1306.4732S}, but tests using
simulated lenses are desirable.

In this paper, we use a new non-parametric strong lens modelling framework -- 
\Glass\ -- to determine what combination of lensing, stellar mass, and
stellar kinematic constraints best constrain the projected mass profile and
shape of a gravitational lens. Some significant work has already proceeded in this direction in the literature.
\citet{2007ApJ...667..645R} were the first to test the recovery of strong
lensing surface density profiles using mock data generated from N-body models
{\bf JR. Is this true?}. They found that when including time delays, circularly
averaged surface density distributions can sufficiently well-recovered to make
interesting comparisons with theoretical models\footnote{Cosmological
parameters, however, are more sensitive. \citet{2007ApJ...667..645R} found that
over-restrictive model assumptions can lead to significant bias on the Hubble
parameter, $H_0$, for example, explaining earlier tensions reported in the
literature \citep[e.g.][]{2002ApJ...578...25K,2002astro.ph..4043K}.}. Since then, several groups
have used N-body mock data to explore the fidelity of strong lensing
reconstructions \citep{2007MNRAS.380.1729L, CITES}. We expand on these earlier
works by considering not only the recovery of the circularly averaged surface density profile
of a lens, but also its shape, higher order measures of the mass recovery (a
pixel-by-pixel comparison), and the utility of wrapping in
constraints from stellar mass estimates and/or stellar kinematics.

This paper is organised as follows. In \secref{sec:theory}, we review the basic
lensing, stellar population synthesis, and Jeans mass modelling theory we will
need. We present some simple back-of-the-envelope calculations that illustrate
what type and quality of data we are likely to need to measure mass profiles
and shapes. In \secref{sec:glass}, we describe the \Glass\ lens inversion
algorithm and its stellar kinematic post-processing module. In
\secref{sec:mockdata}, we describe our mock N-body data. In \secref{sec:results}, we present our
results from applying \Glass\ to these mock data. Finally, in
\secref{sec:conclusions} we present our conclusions.

>>>>

\cite{1981ApJ...244..736Y} were remarkably prescient about the
subsequent development of lens modelling.  First, the introduced the
technique of choosing a parametric form for the lensing mass and then
fitting for the parameters, which is still the most common strategy
\citep[see for example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.
Second, they pointed out the non-uniqueness of lens models, or lensing
degeneracies.  Third, they suggested combining lensing data with
stellar kinematics and X-rays, to reduce the effect of the
degeneracies.  Later work, as well as following up these suggestions,
has introduced some ideas.  Four of these are important in the present
work.
\begin{itemize}
\item In free-form models, there is no specified parametric form for
  the mass distribution.  There are still assumptions (or priors) on
  the mass distribution, such as smoothness or being centrally
  concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms.
  \cite{2006MNRAS.367.1209L} implemented a particular elegant prior,
  which is that the mass distribution must be non-negative and no
  extra images are produced.  Free-form are more commonly used with
  cluster lenses\citep{2009ApJ...690..154S,2013arXiv1304.2393S}, but
  can be used with galaxy lenses as well, where their less restrictive
  assumptions can be important.  For example, in time-delay galaxy
  lenses parametric models led to tensions with the Hubble parameter
  $H_0$ \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}, but
  the less restrictive assumptions of free-form models resolved these
  tensions \citep{2007ApJ...667..645R}.
\item Model ensembles, exploring a diverse range of possible mass
  distributions that nonetheless all fit the data, are a way of
  cambating the non-uniqueness of models.  Model ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are
  more common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.
\item Stellar kinematics has played a role in many modelling
  techniques.  The underlying reason is the virial theorem, which
  relates the line of sight velocity dispersion to the Einstein radius
  as
\begin{equation}
  \frac{\langle v^2_{\rm los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\end{equation}
  with the relation becoming exact for isothermal lenses.  The use of
  two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is
  especially interesting.
\item The stellar mass in a lens can be inferred from photometry and
  compared with the total mass
  \citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.
  Since the inferred stellar mass does depend on the assumed IMF,
  lenses in which stellar mass dominates can be used to derive upper
  bounds on the stellar $M/L$ \cite{2010MNRAS.409L..30F}. Lower bounds
  on stellar $M/L$ have also been claimed \citep{2013MNRAS.428.3183D}
  but these require upper-bound assumptions about the dark-matter
  profile.
\item Testing modelling strategies to see how well they recover
  simulated lenses is increasingly recognized as essential.  Simple
  blind tests have appeared in earlier work \citep[for example, Figure
    2 in][]{2000AJ....119..439W}, but more recently, test against
  dynamically simulated galaxies are favoured
  \citep{2007ApJ...667..645R,2009MNRAS.393.1114B}.
\end{itemize}
There are also more modelling ideas in the literature.  One is to use
X-ray intensity and temperature profiles as a mass constraint
\citep[such as in][]{2013ApJ...765...25N}.  Another idea is to model
multiple lenses simultaneously, with one or more cosmological
parameters variable but shared between the lenses.  This strategy has
been used to constrain $H_0$ from time delays
\citep{2006ApJ...652L...5S,2008ApJ...679...17C,2010ApJ...712.1378P}.
In principle, cosmological parameters could be constrained even from a
single lens.  Some models have attempted this
\citep{2010Sci...329..924J,2013arXiv1306.4732S} starting from a prior
based on the values and uncertainties from the CMB, but tests using
simulated lenses are desirable.

In this paper, we use a new program --- \Glass\ --- to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of a
gravitational lens.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing modules.  \Glass\ is
described in \secref{sec:glass}.  In \secref{sec:theory} before that
we review the elements of lensing theory, stellar population
synthesis, and stellar dynamics we will need.

In \secref{sec:mockdata}, we describe our mock data.

In \secref{sec:results}, we present our results from applying \Glass\ to
these mock data.

\cite{2007ApJ...667..645R} were the first to test the recovery of
strong lensing surface density profiles using mock data generated from
N-body models {\bf JR. Is this true?}. They found that when including
time delays, circularly averaged surface density distributions can
sufficiently well-recovered to make interesting comparisons with
theoretical models Cosmological parameters, however, are more
sensitive. Since then, several groups have used N-body mock data to
explore the fidelity of strong lensing reconstructions
\citep{2007MNRAS.380.1729L}. We expand on these earlier works
by considering not only the recovery of the circularly averaged
surface density profile of a lens, but also its shape, higher order
measures of the mass recovery (a pixel-by-pixel comparison), and the
utility of wrapping in constraints from stellar mass estimates and/or
stellar kinematics.
